---
title: |
  | ECON 4101 Econometrics
  | Reproducibility Term Paper
author: "Pranav Singh"
date: "April 28, 2017"
output: pdf_document
# output: html_document
# output: github_document
---

```{r, include=F}
library(knitr)
# opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=T)
options(scipen=999)

library(ggplot2)
library(gridExtra)
require(data.table)
theme_set(theme_bw())
```

# Background Overview
There are two main competing views of the global market for crude oil:  
  
  1. The price of oil is determined by desired stocks. Shifts in the expectations of forward-looking traders (hedgers and speculators) are reflected in changes in the real price of oil and in changes in oil inventories.
  2. The real price of oil is determined by shocks to the amount of oil coming out of the ground ("flow supply of oil") and the amount of oil being consumed ("flow demand for oil").

Recently, there has been increasing recognition that both elements of price determination matter. In _The Role of Inventories and Speculative Trading in the Global Market for Crude Oil (2013)_, Kilian and Murphy develop a structural vector autoregressive (VAR) model of the global market for crude oil that nests these two theories together and quantifies the effects of shocks to the speculative demand for oil as well as shocks to flow demand and flow supply of oil. The study stresses the importance of oil inventories data for building a model for price of crude oil. The paper attempts to indirectly identify the effects of shifts in expectations (i.e. without explicit measures of expectations) by including changes in oil inventories in their econometric model. They do this because reliable and substantial data on expectation shifts is not readily available, and because how subjective expectations are formed is too complicated of a nonlinear function to be practical in modeling. The model is built off of four main features (measured monthly from 1973:2 to 2009:8):

  1. Percent change in global crude oil production
  2. Index of global real activity in deviations from trend
  3. Real price of oil
  4. Change in above-ground global crude oil inventories
  
The authors assert that the variables are mutually endogenous. They include two years worth of lags in the model. The paper models four shocks as its responses:

  1. Flow supply shock - shock to the flow of crude oil production
  2. Flow demand shock - shock to the demand for crude oil driven by the global business
  3. Speculative demand shock - shock to the demand for above-ground oil inventories arising from forward-looking behavior
  4. Residual shock - captures all structural shocks not otherwise accounted for
  
\pagebreak

# Methods & Procedures

## VAR Methodology
The analysis is based on a dynaic simultaneous equation model in the form of a structural vector autoregression (VAR). Let $y_t$ be a vector of endogenous variables including the percent change in global crude oil production, a measure of global real activity, the real price of crude oil, and the change in global oil inventories above the ground. All data are aggregated monthly. The sample time period is 1973:2 to 2009:8. Seasonal variation is removed by including seasonal dummies in the VAR model. 

## Data
The dry cargo shipping rate index (as developed by Kilian, 2009a) are used as the measure for fluctuations in global real activity. This index is designed to capture shifts in the global use of industrial commodities. Data on global crude oil production are gathered from the *Monthly Energy Review* of the Energy Information Administration (EIA). Oil production is expressed in period-over-period percent changes in the model. The real price of oil is defined as the US refiners' acquisition cost for imported crude oil, as reported by the EIA, extrapolated from 1974:1 back to 1973:1 as in Barsky and Kilian (2002) and deflated by the US CPI. The data for total US crude oil inventories is also provided by the EIA. These data are scaled by the ratio of OECD petroleum stocks over US petroleum stocks, also obtained from the EIA. That scale factor ranges from about 2.23 to 2.59 in the paper's data sample. The resulting proxy for global crude oil inventories is expressed in unit changes rather than percent changes for two main reasons. Firstly, percent changes do not appear to be covariance stationary, whereas the unit changes do. Secondly, expressing in unit changes was deemed necessary for properly computing the oil demand elasticity.

## Structural VAR model for Global Market for Crude Oil
The reduced-form model allows for two years' worth of lags. The corrresponding structural model of the global oil market wass written as:
\[ \beta_0 y_t = \sum_{i=1}^{24} \beta_1 y_{t-i} + \epsilon_t \]
where $\epsilon_t$ is the vector of orthogonal structural innovations and $\beta_i$ denotes the coefficient matrices for $i = 0, \dots, 24$. The seasonal dummies were suppressed for notaional convenience. The vector $\epsilon_t$ consists of our four structural shocks:

1. Flow Supply Shock - incorporates supply disruptions associated with exogenous political events in oil-producing countries as well as unexpected politically motivated supply decisions by OPEC members and other flow supply shocks
2. Flow Demand Shock - incorporates to the demand for crude oil and other industrial commodites that is associated with unexpected fluctuations in the global business cycle
3. Speculative Demand Shock - captures shifts in the demand for above-ground oil inventories arising from forward-looking behavior not otherwise captured by the model
4. Residual Shock - captures idiosyncratic oil demand shocks no otherwise acounted for (e.g. weather shocks, changes in technology or prefernces)



# Part I: Replication of Structural Impulse Responses of Shocks 
```{r, child='Figure1.Rmd'}
```

\pagebreak

# Part II: Replication of Historical Decompositions of Cumulative Effects of Shocks
```{r, child='Figure2.Rmd'}
```

\pagebreak

# Discussion

The paper offers explanations for several historical price shocks. In particular, its conclusions on the 2003-08 oil price shock run counter to the prevailing explanations at the time the paper was written; the paper discounts the suggestions that "peak oil" theory, OPEC supply, or speculation by oil traders were responsible for the oil price increase. Instead, the paper claims there is strong evidence that a booming world economy was the cause.

The paper offers three policy conclusions:

  1. Increased regulation of oil traders will not keep the real price of oil down.
  2. Increased domestic oil production in the U.S. will not lower the real price of oil materially.
  3. Efforts to revive the world economy will cause the real price of oil to recover.

