---
title: |
  | ECON 4101 Econometrics
  | CM08 Homework
author: "Pranav Singh"
date: "March 24, 2017"
output: pdf_document
# output: html_document
# output: github_document
---

```{r, include=F}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=T)
options(scipen=999)

library(ggplot2)
library(gridExtra)
require(data.table)
theme_set(theme_bw())

require(xlsx)
```

```{r}
df <- read.xlsx('../../Data_Woolridge/athlet1.xls', 1)
df <- data.frame(year = df[,1], apps = df[,2], stufac = df[,6], bowl = df[,7])
str(df)
```

```{r}
summary(lm(log(apps) ~ log(stufac) + factor(bowl), df))
```

From the above regression analysis, we find at the 5% significance level, whether or not a university had a bowl game in the previous year has a nearly significant (it's p-value = .0516) impact on admissions ain the current year after accounting for the student-faculty ratio. The elasticity of demand for admissions using the student-faculty ratio is simply that variable's parameter estimate and is equal to -0.25978. That is, holding whether the university had a bowl game the previous year constant, a 1% increase in the student-faculty ratio corresponds to an estimated 0.25978% decrease in the applications for admissions. 
