---
title: |
  | ECON 4101 Econometrics
  | CM23 Homework
author: "Pranav Singh"
date: "April 12, 2017"
output: pdf_document
# output: html_document
# output: github_document
---

```{r, include=F}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=T)
options(scipen=999)

library(ggplot2)
library(gridExtra)
require(data.table)
theme_set(theme_bw())

require(xlsx)
require(lmtest)
require(sandwich)
require(nlme)
```

```{r}
df <- read.csv('../../Data/monop.csv')
str(df)
lay <- rbind(c(1,2), c(3))
g1 <- ggplot(df, aes(x = q, y = tr)) + geom_point() + geom_smooth(method='loess') + scale_x_continuous(breaks=pretty(df$q, n=10)) + scale_y_continuous(breaks=pretty(df$tr, n=10)) + ylab('Total Revenue') + xlab('Quantity')
g2 <- ggplot(df, aes(x = q, y = tc)) + geom_point() + geom_smooth(method='loess') + scale_x_continuous(breaks=pretty(df$q, n=10)) + scale_y_continuous(breaks=pretty(df$tc, n=10)) + ylab('Total Cost') + xlab('Quantity')
g3 <- ggplot(df, aes(x = q, y = tr-tc)) + geom_point() + geom_smooth(method='loess') + scale_x_continuous(breaks=pretty(df$q, n=20)) + scale_y_continuous(breaks=pretty(-5000:10000, n=20)) + ylab('Profit = TR - TC') + xlab('Quantity')
grid.arrange(grobs=list(g1,g2,g3), layout_matrix = lay)
```

Monopolist's Economic Model:
\[
\begin{aligned}
tr &= \beta_1q + \beta_2q^2 \\
tc &= \alpha_1 + \alpha_2q + \alpha_3q^2
\end{aligned}
\]

# Part A
\[
\begin{aligned}
mr &= \frac{d}{dq}\{tr\}  = \beta_1 + 2\beta_2q \\
mc &= \frac{d}{dq}\{tc\} = \alpha_2 + 2\alpha_3q
\end{aligned}
\]

# Part B
\[
\begin{aligned}
& mr & = \; & mc \\
\implies & \beta_1 + 2\beta_2q^* & = \;& \alpha_2 + 2\alpha_3q^* \\
\implies & (2\beta_2 - 2\alpha_3)q^* & = \;& \alpha_2 - \beta_1 \\
\implies & q^* & = \;& \frac{\alpha_2 - \beta_1}{2(\beta_2 - \alpha_3)}
\end{aligned}
\]

# Part C
```{r}
m.tr <- lm(tr ~ 0 + q + I(q^2), df); summary(m.tr)
m.tc <- lm(tc ~ q + I(q^2), df); summary(m.tc)
```
These estimates are appropriate (unbiased) for models that satisfy the first four Gauss-Markov assumptions: (1) the model is linear in parameters, (2) the observations are randomly sampled, (3) the model has zero conditional mean, and (4) there is no perfect collinearity between the model's terms.

```{r}
profitMaximizingQuantity <- function(alpha, beta) {
  (alpha[2] - beta[1]) / (2*(beta[2] - alpha[3]))
}
qstar <- profitMaximizingQuantity(m.tc$coefficients, m.tr$coefficients)
```
From part (b), we see that the OLS estimates suggest the profit maximizing level of output is $q^*$ = `r qstar`.

# Part D
```{r}
qstar <- round(qstar)
pred.tr <- predict(m.tr, data.frame(q=qstar))
pred.tc <- predict(m.tc, data.frame(q=qstar))
pred.profit <- pred.tr - pred.tc
```
The predicted profit is \$`r round(pred.profit, 2)`.

# Part E
```{r}
for(i in 1:5) {
  bg <- bgtest(m.tr, order=i)
  s <- sprintf("Breusch-Godfrey Test: AR(%d) P=%0.4f \n", i, bg$p.value)
  cat(s)
}
for(i in 1:5) {
  bg <- bgtest(m.tc, order=i)
  s <- sprintf("Breusch-Godfrey Test: AR(%d) P=%0.4f \n", i, bg$p.value)
  cat(s)
}
```

# Part F
```{r}
acf(resid(m.tr))
m.tr.ols <- gls(tr ~ 0 + q + I(q^2), df)
m.tr.gls <- gls(tr ~ 0 + q + I(q^2), df, correlation = corARMA(p=1,q=0))
anova(m.tr.ols, m.tr.gls)
summary(m.tr.gls)
confint(m.tr.gls)
```
```{r}
acf(resid(m.tc))
m.tc.ols <- gls(tc ~ q + I(q^2), df)
m.tc.gls <- gls(tc ~ q + I(q^2), df, correlation = corARMA(p=1,q=0))
anova(m.tc.ols, m.tc.gls)
summary(m.tc.gls)
confint(m.tc.gls)
```

# Part G
```{r}
qstar <- profitMaximizingQuantity(m.tc.gls$coefficients, m.tr.gls$coefficients)
```
The profi-maximizing level of output suggested by the GLS model is $q^*$ = `r qstar`.

# Part D
```{r}
qstar <- round(qstar)
pred.tr.gls <- predict(m.tr.gls, data.frame(q=qstar))
pred.tc.gls <- predict(m.tc.gls, data.frame(q=qstar))
pred.profit.gls <- pred.tr.gls - pred.tc.gls
```
The predicted profit is \$`r round(pred.profit.gls, 2)`.

