---
title: |
  | ECON 4101 Econometrics
  | CM15 Homework
author: "Pranav Singh"
date: "March 5, 2017"
output: pdf_document
# output: html_document
# output: github_document
---

```{r, include=F}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=T)
options(scipen=999)

library(ggplot2)
library(gridExtra)
require(data.table)
theme_set(theme_bw())
```

```{r}
data <- fread('cobb_douglas_data.txt')
data[, Year := as.Date(paste0(Year, '-01-01'))]
attach(data)
```

We model the data using a transformation of the Cobb-Douglas production function:
\[ \ln{Q} = \ln{A} + \alpha\ln{L} + \beta\ln{K} \]
where $Q$ = output, $L$ = labor, $K$ = capital, and $A$ = constant.

```{r}
log.Output <- log(Output)
log.Labor <- log(Labor)
log.Capital <- log(Capital)
mod1 <- lm(log.Output ~ log.Labor + log.Capital)
summary(mod1)
confint(mod1)
```
```{r, include=F}
coefs.orig <- coef(mod1)
coefs.orig
```

The assumption of constant returns to scale requires that $\alpha + \beta = 1$. We test this assumption using the following restricted least squares framework:
\[
\begin{aligned}
\text{Original Model: } & \ln{Q} = \ln{A} + \alpha\ln{L} + \beta\ln{K} \\
\text{Restriction: } & \alpha + \beta = 1 \\
\implies & \ln{Q} = \ln{A} + (1-\beta)\ln{L} + \beta\ln{K} \\
\implies & \ln{Q} - \ln{L} = \ln{A} + \beta(\ln{K} - \ln{L})
\end{aligned}
\]
```{r}
log.Output <- log.Output - log.Labor
log.Capital <- log.Capital - log.Labor
mod2 <- lm(log.Output ~ log.Capital)
summary(mod2)
```

```{r}
beta <- mod2$coefficients['log.Capital']
alpha <- 1 - beta
ratio <- Output/(Labor^alpha * Capital^beta)
h <- exp(mod2$coefficients['(Intercept)'])
ggplot() + geom_line(aes(x = Year, y = ratio)) + geom_hline(aes(yintercept = h), linetype = 'dashed')
```


