---
title: |
  | ECON 4101 Econometrics
  | CM15 Homework
author: "Pranav Singh"
date: "March 5, 2017"
output: pdf_document
# output: html_document
# output: github_document
---

```{r, include=F}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=T)
options(scipen=999)

library(ggplot2)
library(gridExtra)
require(data.table)
theme_set(theme_bw())
```

```{r}
data <- fread('cobb_douglas_data.txt')
data[, Year := as.Date(paste0(Year, '-01-01'))]
attach(data)
```

We model the data using a transformation of the Cobb-Douglas production function:
\[ \ln{Q} = \ln{A} + \alpha\ln{L} + \beta\ln{K} \]
where $Q$ = output, $L$ = labor, $K$ = capital, and $A$ = constant.

```{r}
log.Output <- log(Output)
log.Labor <- log(Labor)
log.Capital <- log(Capital)
mod1 <- lm(log.Output ~ log.Labor + log.Capital)
summary(mod1)
confint(mod1)
```

One way to test the hypotheses
\[ 
H_0 \; : \; \alpha + \beta = 1 \\
H_1 \; : \; \alpha + \beta \neq 1 
\]
would be to use the t-statistic:
\[ t_{\hat{\alpha} + \hat{\beta}} = \frac{(\hat{\alpha} + \hat{\beta}) - 1}{\text{se}(\hat{\alpha} + \hat{\beta})} \]
We can get the standard error from the covariance matrix of the estimators of the unrestricted fitted model as follows:
\[ \text{se}(\hat{\alpha} + \hat{\beta}) = \sqrt{\widehat{\text{VAR}}(\hat{\alpha} + \hat{\beta})} 
                                         = \sqrt{\widehat{\text{VAR}}(\hat{\alpha}) + \widehat{\text{VAR}}(\hat{\beta}) + 2\widehat{\text{COV}}(\hat{\alpha}, \hat{\beta})} \]
```{r}
vcov.params <- vcov(mod1)
vcov.params
se <- sqrt(vcov.params['log.Labor','log.Labor'] + vcov.params['log.Capital', 'log.Capital'] + 2*vcov.params['log.Labor', 'log.Capital'])
t.stat <- unname((mod1$coefficients['log.Labor'] + mod1$coefficients['log.Capital'] - 1)/se)
cat('t statistic: ', t.stat)
n <- nrow(data)
k <- 3
t.crit <- qt(.975, n - k)
cat('t critical value: ', t.crit)
pval <- pt(t.stat, df = n - k, lower.tail = F)
cat('p-value: ', pval)
```


```{r, include=F}
coefs.orig <- coef(mod1)
coefs.orig
```
The above test indicates we cannot reject the assumption that $\alpha + \beta = 1$, at least at the 95% confidence level.

Another way to test the assumption of constant returns to scale would be to use a restricted least squares framework as follows:

\[
\begin{aligned}
\text{Original Model: } & \ln{Q} = \ln{A} + \alpha\ln{L} + \beta\ln{K} \\
\text{Restriction: } & \alpha + \beta = 1 \\
\implies & \ln{Q} = \ln{A} + (1-\beta)\ln{L} + \beta\ln{K} \\
\implies & \ln{Q} - \ln{L} = \ln{A} + \beta(\ln{K} - \ln{L})
\end{aligned}
\]
```{r}
log.Output <- log.Output - log.Labor
log.Capital <- log.Capital - log.Labor
mod2 <- lm(log.Output ~ log.Capital)
summary(mod2)
```

We use the F-statistic from the Ramsey RESET Test to compare our restricted and unrestricted models:
\[ F = \frac{(SSE_R - SSE_U)/q}{SSE_U/(n-k-1)}\]
\[ F \sim F_{\alpha,\; q,\; n - k - 1} \]
where $SSE_R$ is for restricted model and $SSE_U$ is for unrestricted model, and $q$ = number of restrictions, and $k$ is the number of terms in the unrestricted model. An `anova` summary of the two models generates our F-statistic value:
```{r}
anova(mod2, mod1)
```
The insignificant p-value here corroborates our earlier finding that we cannot reject the assumption of constant returns to scale. Our findings here match those of Felipe [2005]. It is important to note, as the authors go to great lengths to explain, that these findings should not be interpreted as support for the inferential capabilities of the Cobb-Douglas production function. Rather, the authors offer a convincing argument that "all estimations of aggregate production functions do is to reproduce the distribution income accounting identity."

```{r}
beta <- mod2$coefficients['log.Capital']
alpha <- 1 - beta
ratio <- Output/(Labor^alpha * Capital^beta)
h <- exp(mod2$coefficients['(Intercept)'])
ggplot() + geom_line(aes(x = Year, y = ratio)) + geom_hline(aes(yintercept = h), linetype = 'dashed')
```


